### 2021-01-02 drafts/Response letter.md
Finished the response letter

## Response letter:

January 1st, 2020

Editor-in-Chief
*Ecoscience*

Dear Dr. Hugo Asselin,

We thank the two anonymous reviewers for their thoughtful and useful comments on the previous of our manuscript (TECO-2020-0114) and we are pleased to submit this revised version for potential publication in *Ecoscience*. We have carefully followed the reviewers' suggestions as well as your editorial concerns, in conducting our revision. We now re-analysised the data and nearly rewrite the manuscript to response to reviewers' major concerns, the main points that have been changed in the new version are:

- We have significantly streamlined the manuscript to make it more focused on our main findings now.
- We have re-analysed the data to ensure that the methods and results are more reliable and that the significance levels and error criteria have been chosen in line with international practice.
- We have revised the presentation of the manuscript to make our experimental approach more understandable.

We have also addressed all the other questions and comments raised by the two reviewers. A detailed response to each question and each comment from the reviewers is attached to this cover letter.

We trust that this new version will satisfy both you and the reviewers. Thank you once again for the time that you and reviewers have spent reviewing this MS.

The manuscript includes 1 text (3903 words), 9 figures, 40 references. 9 figure and 1 tables.

We look forward to hearing from you.

Yours sincerely,
Shuang Song (on behalf of the author team)

## Response to Reviewer #1

We have followed all your suggestions in our revision. Below are the reviewer’s previous comments, followed by our responses.

Sincerely,
Shuang Song

---

***[Reviewer #1 Comment #1]*** The ms of Song et al concerns the adaptive capacity of psammophytes to sand burial as a crucial factor for the ecological restoration of coastal dune systems. The aim of the study was to analyze the responses of a typical for the coastal area of South China psammophytic species (Spinifex littoreus) to different sand burial depths and levels. The study is particularly relevant as coastal dunes in China are highly endangered (as many coastal dune systems in the world). The study is also very interesting and well conducted and reports valuable information, which could be a useful tool in the future for a natural and environmentally friendly way for protection and restoration of coastal dune systems in the area. In my opinion, the ms is worth to be published, just minor changes are requested.
***[Response]*** We are grateful to reviewer #1 for the effort reviewing our paper and his/her positive feedback. The information of our work as written by this reviewer is useful and critical. Here below we address the questions and suggestions raised by the reviewer #1.

***[Reviewer #1 Comment #2]*** Line 116: Figure 1. is the third Figure, mentioned in the text, so I think it would be better to renumber it to Figure 3. and renumber the other Figures accordingly.
***[Response]*** Since we re-analyzed our data, we have re-organized manuscript with figures where numbers are well renumbered.

***[Reviewer #1 Comment #3]*** Lines 117-119: Can you provide a reference, that concerns flora and/or vegetation of nebkhas/foredunes in the area?
***[Response]*** We added some new cites in the proper position of ms now (e. g. (Yang et al. 2017, Divyasree & Raju 2019)) to describe the vegetation in this area.

***[Reviewer #1 Comment #4]*** Line 125: How did you compare the age of different plants?
***[Response]*** Thanks for pointing this out. In fact, we only compare them by size (we mis-wrote in the last version.).

***[Reviewer #1 Comment #5]*** Lines 127-129: Did you measure the height of vertical ramets and the length of stolons to know they were all similar with each other before choosing the plants for the study?
***[Response]*** Yes, we started our measurements in the first day of experiment and the we show them in the figures now [**see Figure 5**].

***[Reviewer #1 Comment #6]*** Line 131: What do you mean by "weeds" - all plant species around the selected plants, or particular species, that you consider undesirable, which tend to overgrow or choke out more desirable plants?
***[Response]*** Thanks for the valuable comment, this has been clarified in the revised version of the manuscript:
**[Line 130-131]** All plant species around the selected plants were manually cleared, and...

***[Reviewer #1 Comment #7]*** Lines 269-272: This sentence is not clear to me.
***[Response]*** Thanks for pointing this out, we have deleted this unclear sentence.

***[Reviewer #1 Comment #8]*** Although I'm not a native English speaker, I believe that English should be improved. I wouldn't say it is poor but it needs to be revised. Some examples:
Lines 28-29: Sand dunes occupy a finite area in coastal regions, but they are characterized by providing multiple ecological services, …
Line 57: The tolerance capacity to different sand burial levels is varied among species, …
Lines 142-143: Since the height of vertical ramets of S. littoreus is relatively high, it is usually impossible for them to be completely buried in the sand.
***[Response]*** Thanks for the valuable comment, this has been clarified in the revised version of the manuscript.

### References

Divyasree, M. & A. J. S. Raju, 2019. An Ecological Study Of Reproduction In Spinifex Littoreus (Burm.F.) Merr. (Poaceae), A Dominant Species Of Sand Dune Ecosystem Along The Visakhapatnam Coast, Andhra Pradesh. Journal of Palynology, 55: 35–54.

Yang, X., J. Du, J. Qin, Z. Chen, L. Yang & S. Song, 2017. Diurnal variation characteristics of leaf water potential of Spinifex littoreus on the nebkhas in different succession periods on the coast of Pingtan Island, Fujian Province, China. Chinese Journal of Applied Ecology, 28: 3260.

## Response to Reviewer #2

We have followed all your suggestions in our revision. Below are the reviewer’s previous comments, followed by our responses.

Sincerely,
Shuang Song

---

***[Reviewer #2 Comment #1]***

The authors conducted an experiment to determine how sand burial added at different rates and in different location at a coastal site impact different growth metrics of Spinifex littoreus. The authors found that sand burial had a limited impact on vertical ramet height regardless of amount of sand burial. However, burial did influence stolon growth, such that complete burial increased stolon growth. The authors also found that adventitious root growth was impacted by sand burial. Adventitious roots were present in all burial treatments and their presence increased as burial amount increased.

Overall, I think that most of the manuscript is well written but could benefit from more editing to make the language clearer for the reader. The main concern I have I have with this study is the length of the manuscript, the justification of the methods, and how the statistical analysis was conducted. I will include concerns and comment for each section below.

***[Response]*** Thank you for acknowledging the importance of our research and we have now restructured the manuscript so that these important questions can be clearly answered and discussed in a focused manner.

**Introduction:**

***[Reviewer #2 Comment #2]*** I think the introduction is well done. The authors include adequate information and have clearly shows their understanding of the dune burial literature. Primarily I think the introduction would benefit from a clear hypothesis statement. The objectives of the project are clear, but I think the reader will benefit from a hypothesis which will orient the reader so that they know what the experts (the authors) are expecting from their study.

***[Response]*** Thanks to your positive comments and useful suggestions, we now clearly give our scientific hypothesis: 
**[P3: L94-96]** In this study, we assume that the growths of *S. littoreus* can be facilitated with the increase of sand burial, which is probably the key factor for this native plant to adapt to the intense sand burial caused by frequent typhoon events in growing seasons.

Some other line by line comments:

***[Reviewer #2 Comment #3]*** Line 63-67 - These two sentences present similar results, but the author makes it seem as though they are contradictory – consider revising.
***[Response]*** Thank you for pointing this out, we have now changed them to a paratactic relationship **[P2, L63-67]**:

> Wang (2005) demonstrated that both stem height in the vertical direction and leaf weight of Messerschmidia sibirica were increased with light sand burial, while they all decreased with moderate treatments contrasted with the control. In addition, Zhou et al. (2015a) showed that the growth of Artemisia desterorum was enhanced with moderate sand burial, but inhibited with intense sand burial.

***[Reviewer #2 Comment #4]*** Line 83 – the term nebkhas is a little jargony - readers that are not familiar with types of dunes will not know what this is. I suggest defining more clearly.
**[Response]** Thank you for pointing this out, we now give a simple note **[P2, L82-83]**:

> ...particularly its stolons on nebkhas (a sand dune that forms around vegetation)...

**Methods:**

***[Reviewer #2 Comment #5]*** There are some revisions that could benefit the flow of the methods, mostly with regard to editing sentence structure.

My primary concern with the methods comes from the lack of detail regarding the timeline of when the measurements were made, etc. Specifically, this is an issue in lines 133-139.

I have a concern with the methods of uncovering and re-covering plants to make measurements. I understand the need to uncover the plants to make the measurements the authors needed to make. But by continuously uncovering and re-covering the plants they are complicating the burial treatment (specifically for with regard to gravity signal strength). But if the authors have found another study that uses a similar method, they should cite that paper for justification. If not, they should provide their own justification to explain how this method will not influence plant response to treatments.

***[Response]*** Thanks to your valuable comments, we have now given a more detailed explanation of the measurement process **[P6, L145-155]**:

>During the measurements, the sand burying the plants was carefully removed to one side in a wooden baﬄe, and the height of the vertical ramets, the length of the stolons and adventitious roots were all measured with a tape. Considering the uncovering and re-covering of sand will probably aﬀect the gravity signal of plants to sand burial, we ﬁnished our measurements in minutes to minimize the potential disturbance, and re-covered the plants to the previous sand burial level with same sand. When the experiments were ﬁnished, both above and below ground biomass of plants were collected and separated according to the label on the stolons. Then all the samples were taken back to the laboratory, and the roots, stems and leaves were separated from the stolons, dried in the oven with a temperature of 65◦C for about 24 h until the weight was constant, and then weighed with an analytical balance (accuracy: 0.0001g).

In summary, as our measurements are usually of short duration (a few minutes) and the overall experiment of long duration (20 days), we believe that we can ignore the effect of the short measurement time on the long-term growth of the plants. By the way, such an experimental approach is similar to that of Zhou et al.

**Data analysis:**

***[Reviewer #2 Comment #6]*** I am confused about why the authors used SD as their measure of variance rather than standard error. When comparing and testing differences of treatments/controls you should be using standard error as the proper metric to represent variation.
***[Response]*** Yes you are right, thank you very much for pointing this out, we confused the two concepts before and have now corrected it.

***[Reviewer #2 Comment #7]*** The use of an alpha-value of 0.1 is not typical in Ecological sciences (at least in the U.S.). Is there a justification for why analysis was conducted with an alpha = 0.1? Usually the common practice is setting the critical value at alpha = 0.05.
***[Response]*** Thank you for your helpful suggestions, we have used 0.05 as the p threshold throughout the new version of manuscript to allow comparison with previous studies.

***[Reviewer #2 Comment #8]*** Based on the scientific design the authors are describing, they should probably be analyzing the data with either a two- or three-way ANOVA. It seems as though plot position (windward, leeward, and swale), days since buried, and amount of burial are all factors with different levels that could be interacting and thus impacting plant response. For this reason, a factorial ANOVA should be considered for analysis. Factorial ANOVAs can be conducted in coordination with repeated measures and I think this should be considered. If the authors have considered these data analysis techniques and chose to not analyze the data in this way, they should provide justification for why they chose not to conduct these analyses.
***[Response]*** We think mix-designed repeated measurements ANOVA is the proper method because it can analyze mixed-effects between treatments and time. It was a pity that we had not concentrated on the mixed-effect in the past version of manuscript. However, we have rectified that and give a sketch to differences between the sand-buried groups and the control groups, in the latest version of manuscript **[see the new figure 4]**.

***[Reviewer #2 Comment #9]*** Some specific comments for the methods section:

Line 107 – change to “Monsoons are…”

Line 141-142 – I don’t think this is a complete sentence.

***[Response]*** Thank you for pointing out these issues, we have revised each of them in the new version:

> [P3, L107] Monsoons are the dominant wind...
>
> [L138-140] Since the vertical ramets of S. littoreus is relatively high, it is unlikely for them to be completely buried in a single typhoon event (Figure 1a). The treated S. littoreus, thus, only the stolons of which were buried at diﬀerent depths and levels (see Figure 3 and Table 1).

***[Reviewer #2 Comment #10]*** Line 145-147 - This is confusing. I am unclear whether there are many different burial levels or if there is only a control group and a burial group? Previous sentences make it seem like there are multiple treatments (line 141-142) but this sentence makes it seem there is just high intensity and a control.
***[Response]*** In the previous version, we presented and discussed the results of the experiments in different locations, which were illustrated due to the small sample of plants on non-windward slopes. We now focus only on the windward slopes, where the species grows most vigorously, to focus on responding to our core hypotheses. The other areas (leeward slopes and swales) we have added to the discussion. In the current version we have therefore deleted the sentence.

***[Reviewer #2 Comment #11]*** Line 165-167 - The Car and Multcomp packages likely have citations included in their R documentation. I think R packages should be cited properly. 
***[Response]*** We have re-analysed the data using Python after correcting the method and have included the corresponding references **[L157-161]**:

> All the data were calculated with average according to all the replications with standard error as measurement of variability. The results were analysed with Python3 and package Pingouin (Vallat 2018). For each observed variable, we used repeated measure ANOVA analysis to test if the experimental treatments were veriﬁed due to time, and checked the signiﬁcance (P < 0.05).

**Results:**

***[Reviewer #2 Comment #12]*** There are a lot of results for this paper and I think some for the most important results get lost in the complexity of the results section as a whole. I think this paper will benefit from pairing down the results to the most important results that contribute most to the gaps in dune grass burial literature.
***[Response]*** Thank you for your valuable suggestions, we have now significantly streamlined the results and focused on two main points:

- Inﬂuence of sand burial on the growths of S. littoreus
- Inﬂuence of sand burial on the biomass allocation of S. littoreus

***[Reviewer #2 Comment #13]*** Specific comments in the results section:
Line 213-216 – This section again appears to be results of a two-way ANOVA in which the author is analyzing how burial is impacted by burial based on time period - however, the statistical analysis section does not talk about the author running a factorial ANOVA.
***[Response]*** We have rewritten the results section and now ensure that both the charts and the text show the interaction issues of time correctly in **[Figure 4]**.

***[Reviewer #2 Comment #14]*** Line 223 – Delete “In half sand burial”
***[Response]*** Thanks for this specific useful comment, rectified.

***[Reviewer #2 Comment #15]*** Line 237-240 – This seems a little unclear. It is unclear how important some of these results are to the study. I think this clarity could be improved by incorporating a hypothesis in the introduction.
***[Response]*** We have re-analysed the data and show the results of the biomass allocation in a dedicated section **[see section result Inﬂuence of sand burial on the biomass allocation of S. littoreus]**.

**Discussion:**

***[Reviewer #2 Comment #16]*** The discussion section is also very long, and I think this paper will be better if it were short than the current version and focus on the informative results. The authors dedicate much of their discussion to comparison with other burial studies, which is a good practice in scientific writing. But, most/all the comparisons are between different species. But we know that burial response is species specific. So, I think the discussion dedicates too much space to species-species comparisons of S littoreus and other species studied by other researchers.
***[Response]*** Again, we have followed your useful advice and will now focus our discussion on two points, now:

- Responses of S. littoreus to severe sand burial
- Signiﬁcance of S. littoreus’ adaptation to sand burial

***[Reviewer #2 Comment #17]*** Specific comments for the results section:
Line 247: This is not an assumption. There are many sand burial studies that have shown this definitively.
***[Response]*** Thanks, you are right, that was not an assumption, rectified:

> [L189] Sand burial can reduce the photosynthetic area...

***[Reviewer #2 Comment #18]*** Line 251-254 – This is where the concern of the methods comes in. i.e. How is the intermittent uncovering and recovering of individuals impacting the results of this study?
***[Response]*** We explained after that:

> [L196-200] This is probably due to the growth characteristics of S. littoreus, which displays both vertical and horizontal extension, the height of its ramets ranges from 40 to 60 cm, and it is virtually impossible to bury the entire vertical ramets by the sand sediments relative to its conjoint stolons during a single typhoon event

We believe that this difference is likely to be an important factor in the ability of the species we studied to adapt to the harsh post-typhoon sand-buried environment.

***[Reviewer #2 Comment #19]*** Line 261-262 – This needs to be more scientifically written. The lateral expansion of dune plants may be about "escaping sand burial" but it is a critical mechanism contributing to building dune ridges that protect the length of the coast.
***[Response]*** Thanks to your helpful suggestions, we have changed the presentation to show this "escape effect" in a more scientific way:

> [L203-206] Consequently, the rapid growth of ramets height in S. littoreus is not considered as the priority for the escape of sand burial, but it inclines to increase the stolon length due to its vulnerability to be buried in a typhoon event, which may contribute to the no signiﬁcant impact on the ramets height by artiﬁcial sand burial in our experiment (Maun 1994).

***[Reviewer #2 Comment #20]*** Line 263-267 – Some of this is either inaccurate or unclear. This sentence suggests that Ammophila and Uniola grow in different areas of the beach. But, Ammophila and Uniola grow in similar beach areas. The difference is with growth strategy Uniola does grow more vertically while Ammophila has more rapid vertical expansion. 
***[Response]*** This sentence has been removed in the new edition, as it does not help much with the main scientific hypothesis we are now concerned with.

***[Reviewer #2 Comment #21]*** Line 294-297 – This is an important sentence, but it needs revised. It is very long for a statement that should be very simple.
***[Response]*** We now discuss this in a much more simple way:

> [L266-268] However, if the sand burial rate is over some threshold, the growth of sand dune species will be inhibited (Maun 1996; Shi et al. 2004).

***[Reviewer #2 Comment #22]*** Line 301-304 – This needs revised slightly. Seems like the authors are saying the height of the ramets causes more burial at different successional stages of dune building/stabilization process?
***[Response]*** We have now added new images **[Figure 9]** to show this discrepancy that we found during the course of our survey. This supports the apparent effect of different levels of sand burial on the growth of this species.

***[Reviewer #2 Comment #23]*** Line 318-321 – I am not sure this sentence contributes much to the current study.
***[Response]*** Thanks for point this out, deleted.

***[Reviewer #2 Comment #24]*** Line 323-326 – I feel like this sentence starts talking about one thing but switches to talk about another? I think the issue lies in the organization of the paragraph. Further, if stress signal of gravity plays a factor is burial response. Then wouldn't uncovering and recovering plants impact the stress signal of gravity through the course of the study?
***[Response]*** As explained in the additional section on methods, our measurements are extremely short relative to the growth experiment, and the results of the long-term experiment are interpreted here.

***[Reviewer #2 Comment #25]*** Line 346-349 – Many of the author's sentences are formatted in this way and I think they should be reconstructed to help the reader follow the sequential processes the author is describing.
***[Response]*** Thank you for your helpful advice, we have checked for similar issues in full manuscript.

***[Reviewer #2 Comment #26]*** Line 389 – I have issue with the use of the word “confirmed” here. I think this is not the correct word selection. findings can't be confirmed by an older study. This study would confirm the findings of Maun 1994, not the other way around.
***[Response]*** You are correct, we have removed a similar misrepresentation, thanks a lot.

***[Reviewer #2 Comment #27]*** Line 356 - This was easily the best portion of this paper. I think the authors should focus and highlight this result/discussion in future drafts of this paper.
***[Response]*** Thank you for such constructive comments, we have devoted half of the new manuscript to presenting and discussing the results in relation to this, which we believe makes the new manuscript more relevant

### References

Maun, M. A., 1994. Adaptations enhancing survival and establishment of seedlings on coastal dune systems. Vegetatio, 111: 59–70.

Vallat, R., 2018. Pingouin: statistics in Python. Journal of Open Source Software, 3: 1026.

Zhou, R., Y. Hou, Y. Wang & J. Zuo, 2015a. The physiological adaptation mechanisms of four common desert species in response to desert environments. Acta Ecologica Sinica, 35: 340–349.

Zhou, R., D. Yang, J. Zuo, Y. Wang, Q. Li & D. Hu, 2015b. Analysis on the growth strategy of creeping stem inVitex trifolia var.simplicifoliadapting to sand burial. Acta Ecologica Sinica, 35: 1165–1174.